---
title: "Newhart ADPCA 2019 v1"
author: "Kathryn Newhart"
date: "December 23, 2018"
output: word_document
---
This version of the document includes time to event caluclations. 

# Case studies
To prepare workspace to run case studies:
* Load ADPCA package
* Set parameters 

Steps for each case study:
1. Declare case study paramters
2. Compile & clean raw data
3. Train ADPCA model
4. Test ADPCA model
5. Calculate fault detection statistics
6. Summarize results

Required packages: 
+ ADPCA
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls())
# devtools::install_github("KNewhart/ADPCA")
library(ADPCA)
```

General ADPCA package parameters:
```{r}
# Where results files should be saved
wd <- "C:\\Users\\Kate Newhart\\Dropbox\\Newhart ADPCA Paper\\R"
setwd(wd)
# Alpha value for statistical analysis
alphaN <- 0.01 
# Number of successive OC observations to trigger an alarm
faultsToTriggerAlarm <- 5 
## Data processing parameters##
# TODO: Location of the KNewhart/MP_SBMBR_data github repository
dataLocation <- "C:/Users/Kate Newhart/odrive/Mines/Data/MP_SBMBR_data/"
# Variables from raw data that will be used to define the state of the bioreactor (BR) and the membrane bioreactor (MT, membrane tank). 
stateVarsBR <- c("BIO_1\\CURRENT_PHASE",
                 "BIO_2\\CURRENT_PHASE",
                 "BIO_BLOWER_1\\RUNNING",
                 "BIO_BLOWER_2\\RUNNING")
stateVarsMT <- c("MBR_1\\CURRENT_MODE",
                 "MBR_2\\CURRENT_MODE",
                 "MBR\\CURRENT_FLUX_MODE",
                 "MBR_1\\CURRENT_STATE",
                 "MBR_2\\CURRENT_STATE",
                 "MBR_1_AIR_SCOUR_VALVE\\COMMAND_1",
                 "MBR_2_AIR_SCOUR_VALVE\\COMMAND_STAT")
# Process variables from raw data that will be used for the BR and MT. 
varsBR <- c("BIO_1\\CURRENT_PHASE",
            "BIO_1\\DO\\PROCESS_VALUE",
            "BIO_2\\CURRENT_PHASE",
            "BIO_2\\DO\\PROCESS_VALUE",
            "BIO_BLOWER_1\\FLOW\\PROCESS_VALUE",
            "BIO_BLOWER_1\\RUNNING",
            "BIO_BLOWER_2\\FLOW\\PROCESS_VALUE",
            "BIO_BLOWER_2\\RUNNING",
            "SEWAGE\\FLOW\\PROCESS_VALUE",
            "SEWAGE\\LEVEL\\PROCESS_VALUE",
            "AMBIENT_TEMP\\PROCESS_VALUE",
            "BIO_1\\LEVEL\\PROCESS_VALUE",
            "BIO_1\\TEMPERATURE\\PROCESS_VALUE",
            "BIO_2\\LEVEL\\PROCESS_VALUE",
            "BIO_2\\TEMPERATURE\\PROCESS_VALUE",
            "RAS_TROUGH\\PH\\PROCESS_VALUE",
            "RAS_TROUGH\\TSS\\PROCESS_VALUE",
            "RAS_TROUGH\\TEMPERATURE\\PROCESS_VALUE")
varsMT <- c("MBR_1\\CURRENT_MODE",
            "MBR_1\\PERM_FLOW\\PROCESS_VALUE",
            "MBR_1\\PERM_PRESS\\PROCESS_VALUE",
            "MBR_2\\CURRENT_MODE",
            "MBR_2\\PERM_FLOW\\PROCESS_VALUE",
            "MBR_2\\PERM_PRESS\\PROCESS_VALUE",
            "MBR\\AIR_SCOUR_FLOW\\PROCESS_VALUE_TANK_1",
            "MBR\\AIR_SCOUR_FLOW\\PROCESS_VALUE_TANK_2",
            "MBR\\CURRENT_FLUX_MODE",
            "MBR_1\\CURRENT_STATE",
            "MBR_2\\CURRENT_STATE",
            "MBR_1\\TRANS_PRESS\\PROCESS_VALUE",
            "MBR_2\\TRANS_PRESS\\PROCESS_VALUE",
            "RAS_TROUGH\\TEMPERATURE\\PROCESS_VALUE",
            "MBR\\AIR_SCOUR_PRESSURE\\PROCESS_VALUE_1",
            "MBR\\AIR_SCOUR_PRESSURE\\PROCESS_VALUE_2",
            "MBR_1_AIR_SCOUR_VALVE\\COMMAND_1",
            "MBR_2_AIR_SCOUR_VALVE\\COMMAND_STAT",
            "MBR\\AIR_SCOUR_PRESSURE\\PROCESS_VALUE",
            "RAS_TROUGH\\DO\\PROCESS_VALUE",
            "RAS_TROUGH\\PH\\PROCESS_VALUE",
            "RAS_TROUGH\\TSS\\PROCESS_VALUE",
            "PERMEATE_TANK\\CONDUCTIVITY\\PROCESS_VALUE",
            "PERMEATE_TANK\\LEVEL\\PROCESS_VALUE",
            "BIO_2\\TSS\\PROCESS_VALUE",
            "BIO_1\\TSS\\PROCESS_VALUE",
            "MBR_1\\INF_FLOW\\PROCESS_VALUE",
            "MBR_1\\LEVEL\\PROCESS_VALUE",
            "MBR_2\\INF_FLOW\\PROCESS_VALUE",
            "MBR_2\\LEVEL\\PROCESS_VALUE",
            "PERMEATE_TANK\\TURBIDITY\\PROCESS_VALUE")
```


## Sewage pump: thermal shutdown (Jan. 29, 2018)
* What: Raw sewage pump shut down
* Why: Clog in septic tank influent, causing the sewage level to drop. This exposed the pump, causing it to overheat/shutdown
* When: Jan 29-30
Variables affected: Sewage Level, Bioreactor Level, Bioreactor DO, Bioreactor Blower Flow,
* SB-MBR Alarm: When bioreactor level dropped below 5 ft
* Notes: System was temporarily shut down on Jan 25th
* Results: No faults detected in the membrane sub-system

### Declare ADPCA parameters
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
keyword <- "sewage_pump_shutdown"
## ADPCA parameters ##
# Date to be evaluated with ADPCA model
testingDay <- as.Date("2018 01 29", format = "%Y %m %d")
# Number of days to train on
rollingWindowDays <- 5
# Number of times model is updated 
updateDays <- rollingWindowDays
# Total number of days to be compiled
nDays <- rollingWindowDays + updateDays
```

### Steps 2-4
```{r message=FALSE, warning=FALSE}
source("case_study_sop.R")
```

### Calculate fault detection statistics
```{r message=FALSE}
trueFaultTime <- as.POSIXct("2018-01-29 09:30 MST")
trueAlarmTime <- as.POSIXct("2018-01-29 13:47 MST")
SSAlarmTime <- index(alarmDataSS)[which(alarmDataSS$Alarm != 0)[1]]
BRAlarmTime <- index(alarmDataBR)[which(alarmDataBR$Alarm != 0)[1]]
# Single vs Multistate
ss_alarm <- as.numeric(trueAlarmTime - SSAlarmTime)
ms_alarm <- as.numeric(trueAlarmTime - BRAlarmTime)
# SPE Single
spe_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(1,3))[1]])
# T2 Single
t2_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(2,3))[1]])
# SPE Multi
spe_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(1,3)))[1]])
# T2 Multi
t2_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(2,3)))[1]])

# ADPCA Statistics Ranking
adpca_statistics <- c(spe_ss_alarm, t2_ss_alarm, spe_ms_alarm, t2_ms_alarm)
adpca_statistics_labels <- c("SPE SS", "T2 SS", "SPE MS", "T2 MS")
# adpca_statistics_labels[match(sort(adpca_statistics), adpca_statistics)]

# Summary
summary_statistics <- as.data.frame(adpca_statistics, row.names = adpca_statistics_labels)
colnames(summary_statistics) <- "Sewage Pump"
# knitr::kable(summary_statistics, digits = 1, caption = "Time to detection (hours)")

# Plot everything
source("case_study_plot.R")
```


## RAS Pump: Clog (Jan. 25, 2018)
* What: Clog slowly developed in RAS pump (between bioreactors and membrane tanks)
* When: Decline (appears) to start on January 5th - January 25th
* Why: Hair clog? Foam? Not sure. 
* Variables: MBR Influent Flow
* Results: Flow declined so slowly prior to the event occuring that it was the corrected process behavior that was detected as OC. 

### Declare ADPCA parameters
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
## ADPCA parameters ##
keyword <- "RAS_pump_clog"
setwd("C:\\Users\\Kate Newhart\\Dropbox\\Newhart ADPCA Paper\\R")
# Date to be evaluated with ADPCA model
testingDay <- as.Date("2018 01 25", format = "%Y %m %d")
# Number of days to train on
rollingWindowDays <- 5
# Number of times model is updated 
updateDays <- 2
# Total number of days to be compiled
nDays <- rollingWindowDays + updateDays

```

### Steps 2-4
```{r message=FALSE, warning=FALSE}
source("case_study_sop.R")
```

### Calculate fault detection statistics
```{r}

trueAlarmTime <- as.POSIXct("2018-01-30 04:43:58 MST")
SSAlarmTime <- index(alarmDataSS)[which(alarmDataSS$Alarm != 0)[1]]
BRAlarmTime <- index(alarmDataBR)[which(alarmDataBR$Alarm != 0)[1]]
# Single vs Multistate
ss_alarm <- as.numeric(trueAlarmTime - SSAlarmTime)
ms_alarm <- as.numeric(trueAlarmTime - BRAlarmTime)
# SPE Single
spe_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(1,3))[1]])
# T2 Single
t2_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(2,3))[1]])
# SPE Multi
spe_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(1,3)))[1]])
# T2 Multi
t2_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(2,3)))[1]])

# ADPCA Statistics Ranking
adpca_statistics <- c(spe_ss_alarm, t2_ss_alarm, spe_ms_alarm, t2_ms_alarm)
adpca_statistics_labels <- c("SPE SS", "T2 SS", "SPE MS", "T2 MS")
# adpca_statistics_labels[match(sort(adpca_statistics), adpca_statistics)]

# Summary
summary_statistics <- cbind(summary_statistics, as.data.frame(adpca_statistics, row.names = adpca_statistics_labels))
colnames(summary_statistics)[2] <- "RAS Pump"
# knitr::kable(summary_statistics, digits = 1, caption = "Time to detection (hours)")

# Plot everything
source("case_study_plot.R")

```


## MT Inf: Clog (Dec. 23, 2017)
What: Decline in MBR1 (MT1) influent
When: Dec 23rd
Why: Clog
Variables: MBR1 & MBR2 influent, BIO1 & BIO2 TSS, RAS TSS


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
### Declare ADPCA parameters
keyword <- "MT_inf_clog"
## ADPCA parameters ##
setwd("C:\\Users\\Kate Newhart\\Dropbox\\Newhart ADPCA Paper\\R")
# Date to be evaluated with ADPCA model
testingDay <- as.Date("2017 12 23", format = "%Y %m %d")
# Number of days to train on
rollingWindowDays <- 5
# Number of times model is updated 
updateDays <- rollingWindowDays
# Total number of days to be compiled
nDays <- rollingWindowDays + updateDays

### Steps 2-4
source("case_study_sop.R")
```



### Calculate fault detection statistics
```{r}
trueFaultTime <- as.POSIXct("2017-12-23 8:00 MST")
trueAlarmTime <- as.POSIXct("2017-12-23 16:00 MST")
SSAlarmTime <- index(alarmDataSS)[which(alarmDataSS$Alarm != 0)[1]]
BRAlarmTime <- index(alarmDataBR)[which(alarmDataBR$Alarm != 0)[1]]
# Single vs Multistate
ss_alarm <- as.numeric(trueAlarmTime - SSAlarmTime)
ms_alarm <- as.numeric(trueAlarmTime - BRAlarmTime)
# SPE Single
spe_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(1,3))[1]])
# T2 Single
t2_ss_alarm <- as.numeric(trueAlarmTime - index(alarmDataSS)[which(alarmDataSS$Alarm == c(2,3))[1]])
# SPE Multi
spe_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(1,3)))[1]])
# T2 Multi
t2_ms_alarm <- as.numeric(trueAlarmTime - index(alarmDataBR)[suppressWarnings(which(alarmDataBR$Alarm == c(2,3)))[1]])

# ADPCA Statistics Ranking
adpca_statistics <- c(spe_ss_alarm, t2_ss_alarm, spe_ms_alarm, t2_ms_alarm)
adpca_statistics_labels <- c("SPE SS", "T2 SS", "SPE MS", "T2 MS")
#adpca_statistics_labels[match(sort(adpca_statistics), adpca_statistics)]

# Summary
summary_statistics <- cbind(summary_statistics, as.data.frame(adpca_statistics, row.names = adpca_statistics_labels))
colnames(summary_statistics)[3] <- "MT Inf"
knitr::kable(summary_statistics, digits = 1, caption = "Time to detection (hours)")

# Plot everything
source("case_study_plot.R")
```

